Spectroscopic confirmation of very-wide stellar binaries and large-separation comoving pairs from \textsl{Gaia DR1}

### 2017-10-02 paper/comoving-followup.tex
footnote - uh oh infinite jest is rubbing off on me

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\tmass}{\acronym{2MASS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\logg}{\ensuremath{\ensuremath{\log g}}}
\newcommand{\teff}{\ensuremath{\ensuremath{T_{\rm eff}}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}

@@

% TODO: assumptions:
% - Same velocity dispersion for all ages. Could have an age-dependent
%   dispersion
% -

@@

The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
to be the probability of the magnitude of the 3D separation of two vectors under
the assumption that the two vectors were drawn independently from an isotropic,
3D Gaussian with dispersion $\sigma$.
That is,
\begin{eqnarray}
    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
\end{eqnarray}
where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.

@@

Dispersion ... Nordstrom 2004

>>>>

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\tmass}{\acronym{2MASS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\logg}{\ensuremath{\ensuremath{\log g}}}
\newcommand{\teff}{\ensuremath{\ensuremath{T_{\rm eff}}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
\DeclareMathOperator{\Tr}{Tr}

@@

@@

The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
to be the probability of the magnitude of the 3D separation of two vectors under
the assumption that the two vectors were drawn independently from an isotropic,
3D Gaussian with dispersion $\sigma$.
That is,
\begin{eqnarray}
    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right]
        \label{eq:vel-diff-dist}\\
    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
\end{eqnarray}
where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.

@@

We use an isotropic model for the field velocity distribution with dispersion
set by taking the trace of the local velocity dispersion tensor for young
($<2~{\rm Gyr}$) thin disk stars from \citealt{Nordstrom:2004}, $\mat{V}_{\rm
N04}$, $\sigma_{\rm field} =
\sqrt{\frac{1}{3}\Tr{\left(\mat{V}_{\rm N04}\right)}} \approx
18~\kms$.\footnote{Even though the dispersion in this isotropic model for the
field stars is only $18~\kms$, most randomly paired stars will tend to have 3D
velocity differences $>20~\kms$. The 15th and 85th percentiles of the
distribution of velocity differences (\eqname~\ref{eq:vel-diff-dist}) are
$\approx 23~\kms$ and $\approx 59~\kms$, respectively.}
This is a conservative choice: the velocity dispersion of thin disk stars
increases steadily with age (e.g., \citealt{Aumer:2016}), so assuming that all
of our pairs are young should lead to under-predictions of the probability that
a given pair is truly comoving.
A more robust model that simultaneously models the ages and age distribution of
the comoving pairs would lead to more precise inferences of these probabilities,
but is outside of the scope of this work.

We generate samples in $f$ using \package{emcee} (\citealt{Foreman-Mackey:2013})
with 28 walkers, keeping track of the value of the marginal likelihood under
each mixture model component for each sample.
We run the sampler for 256 burn-in steps from a small ball of initial positions
centered on $f=0.5$, then restart the walkers from the final position and run
for an additional 1024 steps.
With the samples in $f$ and the mixture model component likelihoods, we compute
the per-pair probabilities of belonging to the comoving component (see the final
section of \citealt{Foreman-Mackey:2014}).

% TODO: summary of assumptions:
% - parallax-proper-motion covariances small
% - Same velocity dispersion for all ages. Could have an age-dependent
%   dispersion
% - Isotropic field
